---
title: "Age-1 Pulse Assignments"
author: "Emilie Geissinger"
date: "May 15, 2019"
output:
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(lubridate)
library(knitr)
```

```{r include=FALSE}
# Data set up
# ---- load data -----
age0<-read.csv("../data/data-working/pulse_range_0_final.csv")
age1<-read.csv("../data/data-working/age-1-mixture-dist.csv")
catch.haul<-read.csv("../data/data-working/catch_haul.csv")
fullcount<-read.csv("../data/data-working/newman-catch.csv")
length<-read.csv('../data/data-working/newman-length.csv')
range_final<-read.csv("../data/data-working/pulse_range_age1_final.csv")

# --- check data ----
# age 0
str(age0)
summary(age0)
head(age0)
dim(age0)
names(age0)

age0<-age0%>%
  mutate(cohort=year)
names(age0)

#age 1
str(age1)
dim(age1)
summary(age1)
head(age1)
names(age1)

# full count
summary(fullcount)
str(fullcount)
dim(fullcount)
head(fullcount)
names(fullcount)

trips<-length%>%
  select(year,julian.date,trip,month,day)%>%
  group_by(year,trip)%>%
  filter(julian.date==min(julian.date))%>%
  distinct(trip,year,julian.date,month,day)

# ---- Date manipulation ----

#format date
fullcount$date<-ymd(paste(fullcount$year,fullcount$month,fullcount$day,sep="-"))
age1$date<-ymd(paste(age1$year,age1$month,age1$day,sep="-"))
age1$julian.date<-yday(age1$date)
length$date<-ymd(paste(length$year,length$month,length$day,sep="-"))
length<-length%>%
  mutate(cohort=year-1)%>%
  mutate(date2=date+3)
# add trips to age 0 to add date
age0<-left_join(age0,trips)
age0$date<-ymd(paste(age0$year,age0$month,age0$day, sep="-"))
head(age0)

#mean size for age 0
age0data<-age0%>%
  select(date,cohort,month,age,pulse,min,max)%>%
  mutate(meanSL=(min+max)/2)

#create min and max for age 1
# 2 standard deviations
age1data<-age1%>%
  select(date, cohort, month, age, dummy_pulse,mu,mu.se,sigma,sigma.se)%>%
  filter(mu<300)%>% # get rid of outliers (>300 mm SL)
  rename(meanSL=mu,stdev=sigma,pulse=dummy_pulse)%>%
  mutate(min=meanSL-stdev,max=meanSL+stdev)%>%
  select(-stdev,-sigma.se,-mu.se)

names(age0data)
names(age1data)
growth<-bind_rows(age0data,age1data)
unique(growth$pulse) # number of pulses present
growth<-growth%>%filter(meanSL<250)%>%
  filter(max<250)

# ---- format count data -----
# go from wide format to long format
catch.haul2<-catch.haul%>%
  gather(key=p1,value=count, starts_with('count'))%>%
  gather(key=p2,value=extrap, starts_with('extrap'))%>%
  separate(p1,c("count2","pulse1"))%>%
  separate(p2,c("extrap2","pulse2"))

# make count df and extrap df
count_only<-catch.haul2%>%
  select(-pulse2,-extrap,-extrap2,-count2)%>%
  rename(pulse=pulse1)
extrap_only<-catch.haul2%>%
  select(-pulse1,-count,-count2,-extrap2)%>%
  rename(pulse=pulse2)
head(count_only)
head(extrap_only)
#combine adjusted df to be one complete 
count_long<-left_join(count_only,extrap_only)

# change NAs to 0 (confirmed that NAs are actually 0s)
count_long$count[is.na(count_long$count)]<-0
```
### Pulse Figures
***

#### Age 1 pulses graphed with age 0 grouped by cohort
Age 0 and and age 1 pulses. Point represents mean standard length. Horizontal bars are min and max standard length for age 0's, and +/- 1 standard deviation for age 1's.

```{r echo=FALSE,warning=FALSE}
cohort.graph<-function(growth,na.rm=TRUE, ...){
  growth<-growth%>%filter(cohort!=1995)
  cohort_list<-unique(growth$cohort)
  pdf("mixture-dist.pdf")
  for (i in seq_along(cohort_list)) {
    plot<-ggplot(subset(growth,growth$cohort==cohort_list[i]),
                 aes(x=date,y=meanSL,group=as.factor(cohort),shape=factor(pulse)))+
      geom_point(size=2)+
      geom_errorbar(aes(ymin=min,ymax=max),width=0)+
      theme_bw()+
      ggtitle(paste(cohort_list[i], "Cohort"))+
      xlab("Date")+ylab("Standard length (mm)")+
      scale_x_date(date_breaks="1 month",
                   date_labels="%b-%y")+
      theme(axis.text.x=element_text(angle=45,vjust=0.5))+
      ylim(c(15,250))
    print(plot)
  }
}
cohort.graph(growth)
```

***

#### Age 1 only
These plots show the estimated mean (point) of each pulse with +/- sd. The points represent each measured fish with a day offset of +3.
```{r echo=FALSE, warning=FALSE}
#As a loop
cohort2.graph<-function(growth,length,na.rm=TRUE, ...){
  growth<-growth%>%filter(age==1)
  length<-length%>%filter(age==1)

  
  cohort_list<-unique(growth$cohort)
  
  pdf("mixtures-length.pdf")
  for (i in seq_along(cohort_list)) {
    plot<-ggplot()+
      geom_point(data=subset(growth,growth$cohort==cohort_list[i]),
                 aes(x=date,y=meanSL,group=cohort,shape=as.factor(pulse)),size=2)+
      geom_errorbar(data=subset(growth,growth$cohort==cohort_list[i]),
                    aes(x=date,ymin=min,ymax=max),width=0)+
      geom_jitter(data=subset(length,length$cohort==cohort_list[i]),
                  aes(x=date2,y=mmSL),colour='grey50',size=.25)+
      ylim(20,250)+
      theme_bw()+
      ggtitle(paste(cohort_list[i], "Cohort"))+
      xlab("Date")+ylab("Standard length (mm)")+
      scale_x_date(date_breaks="1 month",
                   date_labels="%b")+
      theme(axis.text.x=element_text(angle=40))
    print(plot)
  }
  #dev.off()
}
cohort2.graph(growth,length)
```

***

#### All pulse assignments
```{r echo=FALSE, warning=FALSE,include=FALSE}
# update pulse assignments for age 1
str(range_final)
summary(range_final)
range_final<-range_final%>%
  filter(!is.na(min), !is.na(max))

pulse_assign1<-data.frame(trip=rep(range_final$trip,range_final$max-range_final$min+1),
                          year=rep(range_final$year,range_final$max-range_final$min+1),
                          pulse=rep(range_final$pulse,range_final$max-range_final$min+1),
                          mmSL=unlist(mapply(seq,range_final$min,range_final$max)))

# Add to age 1 length data
glimpse(length)
glimpse(pulse_assign1)
length1<-length%>%
  filter(age==1)%>%
  select(-pulse)

length_pulse<-left_join(length1,pulse_assign1)
head(length_pulse)
length_pulse$date<-ymd(paste(length_pulse$year,length_pulse$month,length_pulse$day,sep="-"))
length_pulse<-length_pulse%>%
  mutate(cohort=year-1)

length_pulse<-length_pulse%>%
  arrange(desc(cohort))
```
```{r echo=FALSE,warning=FALSE}
# ---- All years as function -----
cohort.graph<-function(length_pulse,na.rm=FALSE, ...){
  
  cohort_list<-rev(unique(length_pulse$cohort))
  pdf("age1-full-colour.pdf")
  for (i in seq_along(cohort_list)) {
    plot<-ggplot(subset(length_pulse,length_pulse$cohort==cohort_list[i]),
                 aes(x=date,y=mmSL,group=cohort,colour=factor(pulse)),shape=factor(pulse))+
      geom_jitter(size=.25)+
      theme_bw()+
      ylim(c(25,250))+
      ggtitle(paste(cohort_list[i], "Cohort"))+
      xlab("Date")+ylab("Standard length (mm)")+
      scale_color_manual(values=c("#E69F00", "#56B4E9", "#009E73", "#F0E442", "#0072B2", "#D55E00", "#CC79A7"))
    print(plot)
  }
}
cohort.graph(length_pulse)
```


### Age 1 Pulse Ranges
```{r echo=FALSE}
kable(range_final, caption = "Table 1. Size range for age 1 pulses grouped by year and trip. Minimum and maximum are expressed as standard length (mm).")
```
